---
title: "FSH 497: Data visualization"
author: "Amelia J. DuVall"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load libraries
library(here)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(ggridges)
library(viridis)
library(patchwork)
library(gridExtra)
library(grid)
```

<br>

This is version `r paste0('0.', format(Sys.time(), '%y.%m.%d'))`

# Data description

The data come from the [{FSAdata}](https://github.com/droglenc/FSAdata) package, which was created by [Derek Ogle](https://derekogle.com/) as part of his book [*Introductory Fisheries Analysis with R*](http://derekogle.com/IFAR/). The dataset contains information on the ages (subsample), and length and mass (all fish) for male and female Siscowet Lake Trout captured at four locations in Michigan waters of Lake Superior.

## Format

A data frame with 780 observations on the following 8 variables.

1) `locID`: Locations (`Blind Sucker`, `Deer Park`, `Grand Marais`, `Little Lake Harbor`) 

2) `pnldep`: Depth of gillnet panel in which the fish was captured  

3) `mesh`: Gillnet stretch mesh measure  

4) `fishID`: Unique fish identification code  

5) `sex`: Sex (`F` = female; `M` = male)  

6) `age`: Assigned ages (years)  

7) `len`: Total length (mm)  

8) `wgt`: Weight (g)  

```{r readin, warning = FALSE, include = FALSE}
df <- read_csv(here("data", "siscowet.csv"))
head(df)
dim(df)
table(df$locID, useNA = "always")
table(df$sex, useNA = "always")
table(df$age, useNA = "always")
table(df$wgt, useNA = "always")
```

***

# Explanatory figures
```{r explanatory, warning = FALSE, echo = FALSE}
hist(df$pnldep)

hist(df$len)

ggplot(data = df, mapping = aes(x = pnldep, y = len, color = sex)) + 
  geom_point() +
  facet_wrap(~locID) 
```

# Expository figure
### Number 1
```{r expository, warning = FALSE, echo = FALSE}
dfv2 <- df %>%
  filter(!is.na(sex), 
         wgt <10000) # outlier?

max(dfv2$wgt, na.rm = TRUE)
max(dfv2$len, na.rm = TRUE)

ggplot(data = dfv2, mapping = aes(x = len, y = wgt, color = sex)) + 
  geom_point() +
  xlab("Total length (mm)") + ylab("Weight (g)") +
  # geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~locID, nrow = 4) +
  scale_color_viridis_d(option = "cividis") +
  theme_minimal() +
  scale_y_continuous(expand = c(0, 0),
                     limits=c(0, 5000),
                     breaks = c(0, 2500, 5000)) +
  scale_x_continuous(expand = c(0, 0), limits = c(0, 800), 
                     breaks = c(0, 200, 400, 600, 800)) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  geom_jitter()
```

### Number 2
```{r ex2, warning = FALSE, echo = FALSE}
p1 <- ggplot(dfv2) +
  geom_density_ridges(aes(x = len, y = locID, fill = locID)) +
  theme_cowplot() + 
  theme(legend.position = "none") +
  xlab("Total length (mm)") + ylab("Location") +
  scale_fill_viridis_d(option = "cividis")
p2 <- ggplot(dfv2) +
  geom_density_ridges(aes(x = wgt, y = locID, fill = locID)) +
  theme_cowplot() + 
  theme(legend.position = "none") +
  xlab("Weight(g)") + ylab("Location") +
  scale_fill_viridis_d(option = "cividis")
p3 <- ggplot(dfv2) +
  geom_density_ridges(aes(x = pnldep, y = locID, fill = locID)) +
  theme_cowplot() + 
  theme(legend.position = "none") +
  xlab("Depth of gillnet panel") + ylab("Location") +
  scale_fill_viridis_d(option = "cividis")
grid.arrange(p1, p2, p3)
```


